---
title: "Methods 2 Class 1"
author: "Victoria Lowe"
date: "2/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,ggplot2,ggfortify,dplyr,stringr,substr,nchar,)
setwd(".")
```

```{r exercise 1.2 A}
#Figure 1.1b shows data corresponding to the fitted line y = 46.3 + 3.0x with residual standard deviation 3.9, and values of x ranging roughly from 0 to 4%.

#A. Sketch hypothetical data with the same range of x but corresponding to the line 
#   y = 30 + 10x with residual standard deviation 3.9.
x1 <- rnorm(1000,mean=2,sd=0.6)
x1
hist(x1)

error<-rnorm(1000,mean=0,sd=3.9)
hist(error)

#Fit new data into model to find corresponding y-values
y1=30+10*x1+error

#Making a dataframe with x and y-values
df_1<-data.frame(x1,y1)

Illustration_1<-lm(y1~x1,data = df_1)
summary(Illustration_1)
autoplot(lm(y1 ~ x1, data = df_1))

```

```{r exercise 1.2 B}
#B. Sketch hypothetical data with the same range of x but corresponding to the line y = 30 + 10x with residual standard deviation 10.

x2<- rnorm(900,mean = 2,sd=0.6)
x2
hist(x2)

error_2<-rnorm(900,mean = 0,sd=10)

#Fitting the data into a model to get the corresponding y-values
y2<-30+10*x2+error_2

#Making a dataframe with x and y-values
df_2<-data.frame(x2,y2)

#Getting it into lm and visualizing the result
Illustration_2<-lm(y1~x1,data = df_1)
summary(Illustration_2)
autoplot(lm(y2 ~ x2, data = df_2))
```

```{r Own dataset}
#Loading in the dataset:
df_epilepsy<-read_csv("epilepsy.csv")
```
About the data set:
Treatment: factor with levels placebo progabide indicating whether the anti-epilepsy drug Progabide has been applied or not, column-type CATEGORICAL COVARIATE.

Base: number of epileptic attacks recorded during 8 week period prior to randomization, column-type CONTINUOUS COVARIATE.

Age: age of the patients, column-type CONTINUOUS COVARIATE.

Seizure_rate: number of epilepsy attacks patients have during the follow-up period, column-type OBSERVATION.

Period: measurement period, column-type TIME.

Subject: patient identification number, column-type ID.

```{r Own dataset: descriptive statistics}
#Use a descriptive statistics and visual investigation to give a preliminary look into the data. (Use some of the things from chapter 2) 

head(df_epilepsy,10)

#Checking out the numbers of rows and columns
nrow(df_epilepsy)
ncol(df_epilepsy)

#Checking the column names
colnames(df_epilepsy)

#Checking summary statistics
summary(df_epilepsy)

#Checking the structure of the data frame
str(df_epilepsy)
```

```{r Visual investigation}
#The basic plots, scatter plot, histogram, QQ-plot

#Scatterplot over seizure.rate and period
df_epilepsy %>% 
  ggplot()+
  aes(x=period,y=seizure.rate,color=treatment)+
  geom_jitter(width=.4,height=.4)+
  geom_point()+
  ggtitle("Scatterplot")+
  labs(x="period",y="Serizure rate")

#Histogram over seizure.rate:
df_epilepsy %>% 
  ggplot()+
  aes(x=seizure.rate,fill="purple")+
  geom_histogram(binwidth = 5)+
  theme_minimal()+
  ggtitle("Histogram over seizure rate")

#QQ-plot over seizure.rate:
qqnorm(df_epilepsy$seizure.rate, main="QQ-plot over seizure rate")
qqline(df_epilepsy$seizure.rate)

#log-transformed QQ-plot - the log-transformation does not work, it says that it needs "finite ylim"
qqnorm(log(df_epilepsy$seizure.rate), main="QQ-plot over log-transformed seizure rate")
qqline(log(df_epilepsy$seizure.rate))

#Histogram over base:
df_epilepsy %>% 
  ggplot()+
  aes(x=base,fill="blue")+
  geom_histogram(binwidth = 5)+
  theme_minimal()+
  ggtitle("Histogram over base")

#Histogram over log-transformed base:
df_epilepsy %>% 
  ggplot()+
  aes(x=log(base),fill="purple")+
  geom_histogram(binwidth = 0.1)+
  theme_minimal()+
  ggtitle("Histogram over log-transformed base")

#QQ-plot over base:
qqnorm(df_epilepsy$base, main="QQ-plot over base")
qqline(df_epilepsy$base)

#QQ-plot over log-transformed base:
qqnorm(log(df_epilepsy$base), main="QQ-plot over base")
qqline(log(df_epilepsy$base))

#Historgram over age:
df_epilepsy %>% 
  ggplot()+
  aes(x=age,fill="purple")+
  geom_histogram(binwidth = 1)+
  theme_minimal()+
  ggtitle("Histogram over age")
```


```{r Models}
#Create several models and discuss what each model reliably can infer anything about. Mixed effect, which random effects? Logistic regression? Interaction effects? Should we dummy code some variables?  
#Making treatment into 1 (placebo) and 2(progabide)
df_epilepsy$treatment<-as.factor(df_epilepsy$treatment)
df_epilepsy$treatment<-as.numeric(df_epilepsy$treatment)
df_epilepsy$treatment<-as.factor(df_epilepsy$treatment)
levels(df_epilepsy$treatment)

#Logistic regression:
regression_1<-glm(treatment ~ seizure.rate, family = binomial, data = df_epilepsy) 
summary(regression_1)

#Mixed effect model with random intercepts:
intercept_model <- lmerTest::lmer(seizure.rate ~ treatment + (1|subject), 
                                  data = df_epilepsy, REML = F)
summary(intercept_model)

#Mixed effect model with random slopes and intercepts:
int_and_slope_model <- lmerTest::lmer(seizure.rate ~ treatment + (1+base|subject),
                                      data = df_epilepsy, REML = F)
summary(int_and_slope_model)

#Interaction effect:
interaction_effect<-lm(seizure.rate ~ treatment*base,data = df_epilepsy)
summary(interaction_effect)


```

```{r Model comparison}
#Do model comparison.

#comparing mixed-effect models:
anova(intercept_model,int_and_slope_model)

#comparing all the models:
anova(intercept_model, int_and_slope_model, interaction_effect)

#calculating R-squared
MuMIn::r.squaredGLMM(intercept_model)
MuMIn::r.squaredGLMM(int_and_slope_model)
MuMIn::r.squaredGLMM(interaction_effect)

PseudoR2(regression_1,c('McFadden'))

```

A short summary of the model (int_and_slope_model):
lmerTest in R is used to produce a linear mixed-effects model. The model has seizure rate as outcome variable and treatment as predictor value. As random effects, the model has random intercepts for subject to take the individual base lines into account as well as by-base random slopes. 
The outcome of the model shows that seizure rate is not significantly modulated by treatment (p > 0.05). When going from the placebo treatment to the progabide treatment, the seizure rate decreases with 0.5964 (p-value = 0.557), so even though the effect is not statistically significant it does approach significance. The whole model accounts for 74.5% of variance (R2 conditional = 0.745, and the fixed effects account for 0,075 % (R2 marginal = 0.00075). 

```{r Exercise 2.3}
#Data processing: Go to the folder Names and make a graph similar to Figure 2.8, but for girls.
#hints: Summarise_at() , str_sub(), pivot_longer() 

#Loading in the data:
All_names_df<-read_csv("allnames_clean.csv")

#Sorting the df so that there is only girl names
Girl_names_df<-All_names_df %>% 
  filter(sex=="F")

#Converting the names into vectors:
girl <- as.vector(Girl_names_df$name)
girl

#Getting the last letter of each name
RIGHT=function(Girl_names_df,n){
  substring(Girl_names_df,nchar(Girl_names_df)-n+1)
}

last_letters<-RIGHT(girl,1)
last_letters

#Making the vectors into a data frame:
last_letter_df<-data.frame(last_letters)

#Adding that column onto the "Girl_names_df"
Girl_names_df["Last_letter"]=last_letter_df["last_letters"]

#Repositioning the column
Girl_names_df <- Girl_names_df %>% 
  relocate(Last_letter, .after = name)

#Summarizing each year:
Girl_names_df<-Girl_names_df %>% 
  group_by(Last_letter) %>% 
  summarise_at(vars(X1880:X2010),funs(sum))
Girl_names_df

#Getting the percentages throughout the years for each letter (using pivot longer which manipulates the df so that there are many rows instead of many columns)
Girl_names_df_rows <- Girl_names_df %>% 
  pivot_longer(!Last_letter,names_to = "Year", values_to = "Percentage")

#Converting to percentage
Girl_names_df_rows <- Girl_names_df_rows %>% 
  group_by(Year) %>% 
  mutate(sum_by_year=sum(Percentage))
Girl_names_df_rows

Girl_names_df_rows <- Girl_names_df_rows %>% 
  mutate(Percentage=(Percentage/sum_by_year)*100)
Girl_names_df_rows
  
round(Girl_names_df_rows$Percentage,digits = 2) #can't get it into the dataframe, as it won't "find" Percentage when added to the section above

#Getting it into a graph
Girl_names_df_rows %>% 
  ggplot()+
  aes(x = Year, y = Percentage, group = Last_letter, colour = Last_letter) + 
  geom_line()+
  ggtitle("Last letters of Girls' names")+
  labs(x="Year",y="Percentage of all Girls' name that year")
```
